\sysfull

### 2018-04-16 paper/hwswsetup.tex

Experiments on the intermittently-powered platform used the
WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
%
The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered the WISP using an RF signal generator emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with the WISP antenna, and
no objects obstructed the path.
%
We affixed the WISP with a paper harness at the edge of a table, so that the
WISP was parallel to the table surface at a height of 10\,cm.
%
For distance-controlled experiments we positioned the signal generator
antenna at a fixed distance from the WISP. The distances were $d=\{15, 30,
50\}$\,cm for the main experiments and 60\,cm for the task downscaling
experiments.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
the interval between edges in the signal.

@@

We evaluated \sys using six benchmarks summarized in
Table~\ref{table:benchmark_table}. %
Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) are from prior
work~\cite{chain,alpaca} and three benchmarks (\textbf{dijkstra},
\textbf{fft}, \textbf{sort}) are new implementations of applications
from the MiBench suite~\cite{mibench,hicks_mibench2_2016}. The
\textbf{fft} implementation is based on the Accelerated FFT library from
TI~\cite{ti-fft}.
%
\todo{add reference}{Przemek}
%
All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.
%
The source code for all benchmarks will be released at \cite{coala_website}.

>>>>

Experiments on the intermittently-powered platform used the
WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
%
The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered the WISP using an RF signal generator emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with the WISP antenna, and
no objects obstructed the path.
%
We affixed the WISP with a paper harness at the edge of a table, so that the
WISP was parallel to the table surface at a height of 10\,cm.
%
For distance-controlled experiments we positioned the signal generator
antenna at a fixed distance from the WISP. The distances were $d=\{15, 30,
50\}$\,cm for the main experiments and 60\,cm for the task downscaling
experiments.
%
We chose relatively shorter distances than in a deployment scenario to reduce
charge time and run experiments faster, after having confirmed ourselves a
result also reported in prior work~\cite{alpaca} that the amount of energy
available in one charge cycle --- the sum of stored energy and incoming energy
during execution -- is the same at distances beyond 50~cm where incoming energy
becomes negligible.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
the interval between edges in the signal.

@@

We evaluated \sys using six benchmarks summarized in
Table~\ref{table:benchmark_table}. %
Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) are from prior
work~\cite{chain,alpaca} and three benchmarks (\textbf{dijkstra},
\textbf{fft}, \textbf{sort}) are new implementations of applications
from the MiBench suite~\cite{mibench,hicks_mibench2_2016}. The
\textbf{fft} implementation is based on the Accelerated FFT library from
TI~\cite{ti_dsp}. All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.
%
The source code for all benchmarks will be released at \cite{coala_website}.

### 2018-04-16 paper/implementation.tex

\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserve performance by enabling an incremental commit mechanism. When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer. This is what we call \emph{intermediate commit}: it has to be completed in one shot, and if power fails during this phase the application will have to resume from the beginning of the last coalesced task. When the intermediate commit is done, a flag can be atomically raised to signify that a valid, persistent copy of dirty pages is available in non-volatile memory. From this point on, the commit process can be resumed even if power fails.
To understand the last phase, referred to as \emph{ultimate commit}, it has to be mentioned that store and shadow buffers are virtual. By default, \sys allocates two memory banks for non-volatile pages, $M_0$ and $M_1$, and all the pages have their initial store copy in $M_0$. A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $M_x$, will end up having its store location in $M_x$ after the ultimate commit. In other words, during the ultimate commit dirty pages are not physically moved, instead their store and shadow locations are swapped. This trick makes \sys save some energy during this phase, avoiding redundant page moves. As for page swaps, the intermediate commit makes use of the on-board DMA.

>>>>

\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two fold: \emph{ensure consistency} by using atomic operations, and \emph{preserve performance} by enabling an incremental commit mechanism. When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer. This is what we call \emph{intermediate commit}: it has to be completed in one shot, and if power fails during this phase the application will have to resume from the beginning of the last coalesced task. When the intermediate commit is done, a flag can be atomically raised to signify that a valid, persistent copy of dirty pages is available in non-volatile memory. From this point on, the commit process can be resumed even if power fails.
To understand the last phase, referred to as \emph{ultimate commit}, it has to be mentioned that store and shadow buffers are virtual. By default, \sys allocates two memory banks for non-volatile pages, $M_0$ and $M_1$, and all the pages have their initial store copy in $M_0$. A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $M_x$, will end up having its store location in $M_x$ after the ultimate commit. In other words, during the ultimate commit dirty pages are not physically moved, instead their store and shadow locations are swapped. This trick makes \sys save some energy during this phase, avoiding redundant page moves. As for page swaps, the intermediate commit makes use of the on-board DMA.
